---
title: "STK01_report"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
---

## Quarto

```{r}
library(tidyverse)
library(patchwork)

load("STK01_all_data.RData")
```

## Design of the experiment


| Stage 1 | Stage 2  | Stage 3  |
|:--------|:---------|:---------|
| AX-o1   | AX-o1/o2 | DX-o1    |
| AY-o1   | AY-o1    | CY-o2    |
| BX-o2   | BX-o2/o1 | AB-o1/o2 |
| BY-o2   | BY-o2    |          |
| C-o1    | C-o1     |          |
| D-o2    | D-o2     |          |
: STK01 design {tbl-colwidths="[20,20,20]"}

### accuracy of participants in phase 1

```{r}

# create variable that reflects order of tt
all_data <- 
  all_data %>% 
  group_by(pNum, phase, tt) %>% 
  mutate(tt_order = 1:n(), .after = tt)

stg1_acc <- 
  all_data %>% 
  filter(block %in% c(7:12)) %>% # second half of stage 1
  group_by(pNum) %>% 
  summarise(mean_stg1_acc = mean(accuracy, na.rm = TRUE))

CD_perf <- 
  all_data %>% 
  filter(phase == 2,
         tt %in% c("C", "D")) %>% 
  group_by(pNum) %>% 
  summarise(mean_CD_Acc = mean(accuracy, na.rm = TRUE))


perf_check <-  left_join(stg1_acc, CD_perf, by = "pNum")

weak_pNum <- 
  perf_check %>% 
  filter(mean_stg1_acc < .8 | mean_CD_Acc < .8) %>% 
  select(pNum) %>% 
  pull()
  


```

### accuracy by phase

```{r}
acc_by_phase <- 
all_data %>% 
  group_by(pNum, phase) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE))
  
acc_by_phase %>% 
  ggplot(aes(x = phase, y = meanAcc)) +
  geom_violin() +
  geom_jitter(width = .05, height = 0)


```

### accuracy by phase and trial type

```{r}

acc_by_tt <- 
all_data %>% 
  filter(! pNum %in% weak_pNum) %>% 
  group_by(pNum, tt, phase) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE))
  
acc_by_tt %>% 
  ggplot(aes(x = tt, y = meanAcc)) +
  geom_violin(aes(fill = tt)) +
  geom_jitter(width = .2, height = 0) +
  facet_grid(~phase) +
  theme_light(base_size = 24)


```




### plotting means across blocks

```{r}

acc_by_block <- 
  all_data %>% 
  group_by(phase, tt, tt_order) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE))

# # Stage 1
# acc_by_block %>% 
#   filter(phase == 1) %>% 
#   ggplot(aes(x = tt_order, y = meanAcc, colour = tt)) +
#   geom_point(size = 4) +
#   geom_line(size = 1)


# Stage 3
acc_by_block %>% 
  filter(phase == 3) %>% 
  ggplot(aes(x = tt_order, y = meanAcc, colour = tt)) +
  geom_point(size = 4) +
  geom_line(size = 1) +
  facet_wrap(~tt)  +
  theme_light()

# Ps removed
N_60_plot <-
  all_data %>% 
  group_by(phase, tt, tt_order) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE)) %>% 
  filter(phase == 3, 
         tt %in% c("CY", "DX")) %>% 
  ggplot(aes(x = tt_order, y = meanAcc, colour = tt)) +
  geom_point(size = 4) +
  geom_line(size = 1) +
  theme_light(base_size = 24) +
  labs(title = "Full sample: N = 60")

N_29_plot <- 
  all_data %>% 
  filter(! pNum %in% weak_pNum) %>% 
  group_by(phase, tt, tt_order) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE)) %>% 
  filter(phase == 3, 
         tt %in% c("CY", "DX")) %>% 
  ggplot(aes(x = tt_order, y = meanAcc, colour = tt)) +
  geom_point(size = 4) +
  geom_line(size = 1) +
  theme_light(base_size = 24) +
  labs(title = "Best Ps: N = 29")

N_60_plot/N_29_plot

```


```{r}
# stats on stage 3

stg3_CX_DY <- 
  all_data %>%
  #filter(! pNum %in% weak_pNum) %>%
  filter(phase == 3,
         tt_order <5,
         tt %in% c("CY", "DX")) %>% 
  group_by(pNum, tt) %>% 
  summarise(meanAcc = mean(accuracy, na.rm = TRUE))

t.test(meanAcc ~ tt, data = stg3_CX_DY, paired = TRUE)


```



## Possible designs for Experiment 2

ver 2.1

| Stage 1 | Stage 2  | Stage 3  |
|:--------|:---------|:---------|
| AX-o1   | AX-o1/o2 | XA-o1    |
| AY-o1   | AY-o1    | XB-o1    |
| BX-o2   | BX-o2/o1 | YA-o2    |
| BY-o2   | BY-o2    | YB-o2    |
        |
: STK02 design {tbl-colwidths="[20,20,20]"}

ver 2.2

| Stage 1 | Stage 2  | Stage 3  |
|:--------|:---------|:---------|
| AX-o1   | AX-o1/o2 | XA-o1    |
| AY-o1   | AY-o1/o2 | XB-o1    |
| BX-o2   | BX-o2    | YA-o2    |
| BY-o2   | BY-o2    | YB-o2    |
        |
: STK02 design {tbl-colwidths="[20,20,20]"}
